## Chp 2 Definitions and Basic Properties

```{r nelsen_chp2_pck}
# load packages
libs<-c("copBasic","lattice","rgl","magrittr")
invisible(lapply(libs, library, character.only = TRUE))
# use lattice for wireframe
# use rgl for persp3d, etc
```

Consider random vars $X$ and $Y$ with distributions functions $F(x)=P(X \le x)$ and $G(y)=P(Y \le y)$ and joint distribution $H(x,y)=P(X \le x, Y \le y)$.

For each pair of $(x,y) \in \mathbb{R} \times \mathbb{R}$, we can associate three numbers
$F(x)$, $G(y)$ and $H(x,y)$ (each n interval [0,1]).

$(x,y) \rightarrow (F(x),G(y)) \in [0,1] \times [0,1] \rightarrow H(x,y) \in [0,1]$

The correspondence that assigns the value of the joint df to each ordered pair of values of the individual dfs is a function called a *copula*.

Need to generalized notion of "nondecreasing" to multivariate functions

### Preliminaries

A "2-increasing" function is a two-dimensional analog of a nondecreasing function of one variable.

Notation:
$$
\mathbf{R}: \text{ the real line } (-\infty,\infty)\\
\overline{\mathbf{R}}: \text{ the extended real line } [-\infty,\infty]\\
\overline{\mathbf{R}}^2: \text{ the extended real plane } \overline{\mathbf{R}} \times \overline{\mathbf{R}}\\
B = [x_1,x_2] \times [y_1,y_2]: \text{a rectangle in } \overline{\mathbf{R}}^2 \text{ with vertices } (x_1,y_1),(x_1,y_2),(x_2,y_1),(x_2,y_2)\\
\mathbf{I}^2 = \mathbf{I} \times \mathbf{I} = [0,1] \times [0,1]\\
\text{2-place real function }H: \text{function whose domain, } DomH \subset \overline{\mathbf{R}}^2,\\ \text{ and whose range, } RanH \subset \mathbf{R}
$$

Definition: Let $S_1$ and $S_2$ be nonempty subsets of $\overline{\mathbf{R}}$ and $H$ be a 2-place real function s.t. $DomH =  S_1 \times S_2 $. Let $B = [x_1,x_2] \times [y_1,y_2]$ be a rectangle with verticesin $DomH$. The $H-volume$ of $B$ is:
$$
V_{H}(B)=H(x_2,y_2)-H(x_2,y_1)-H(x_1,y_2)+H(x_1,y_1)
$$
Definition: A 2-place real function $H$ is $2-increasing$ (aka quasi-monotone) if $V_H(B) \ge 0$ for all rectangles $B$ whose vertices lie in $DomH$.

"H is 2-increasing" $\not\Leftrightarrow$ "H is nondecreasingin each argument"

#### Ex 2.1 

```{r ex2-1, cache=TRUE}
#H volume generator function 
mk_Hvol<-function(fun=NULL){
  
  function(x,y){
    fun(x[2],y[2])-fun(x[2],y[1])-fun(x[1],y[2])+fun(x[1],y[1])
  }
  
}

# make vectorized function
H1<-function(x,y) mapply(function(x,y) max(x,y),x,y)

# create H-volume function
H1_vol<-mk_Hvol(fun=H1)
```

Let $H(x,y)=\max(x,y)$ defined on $\mathbf{I}^2$. H is a nondecreasing function of  $x$ and $y$, but the H-volume is $V_{H}(\mathbf{I}^2)=$`r H1_vol(c(0,1),c(0,1))`, so this function is *not* 2-increasing. See plots \@ref(fig:ex2-1plta), \@ref(fig:ex2-1pltb), and \@ref(fig:ex2-1pltc)

```{r ex2-1plta, cache=TRUE, fig.cap="Contour plot", fig.align="center"}
# for plots
x<-y<-seq(0,1,length=50)
z1<-outer(x,y,FUN="H1") #same as outer(x,y,FUN="pmax")

# color scheme
num_col<-100
#color <- heat.colors(num_col)
color <- rev(rainbow(num_col, start = 0/6, end = 4/6))

contour(x,y,z1)
```

```{r ex2-1pltb, cache=TRUE, fig.cap="Color image", fig.align="center"}
image(x,y,z1,col=color)
```

```{r ex2-1pltc, cache=TRUE, fig.cap="Interactive", fig.align="center"}
zcol <- cut(z1, num_col)
plotid1<-persp3d(x,y,z1, col=color[zcol])

fn<-spin3d(axis = c(0, 0, 1), rpm = 6) # spin function

rglwidget(elementId = "plot3drgl1") %>% playwidget(par3dinterpControl(fn, 0, 10, steps=25),
       step = 0.1, loop = TRUE, rate = 0.75)
```

#### Ex 2.2

```{r ex2-2, cache=TRUE}
# make vectorized function
H2<-function(x,y) mapply(function(x,y) (2*x-1)*(2*y-1), x,y)

# create H-volume function
H2_vol<-mk_Hvol(fun=H2)
```

Let $H(x,y)=(2x-1)(2y-1)$ defined on $\mathbf{I}^2$. $H$ is 2-increasing since $V_{H}(\mathbf{I}^2)=$`r H2_vol(c(0,1),c(0,1))`, but is a decreasing function of $x$ for each $y \in (0,1/2)$ and a decreasing function of $y$ for each $x\in (0,1/2)$. See plots \@ref(fig:ex2-2plta), \@ref(fig:ex2-2pltb), and \@ref(fig:ex2-2pltc)

```{r ex2-2plta, cache=TRUE, fig.pos="center", fig.cap="Contour plot"}
z2<-outer(x,y,FUN="H2")
contour(x,y,z2)
```

```{r ex2-2pltb, cache=TRUE, fig.pos="center", fig.cap="Color Image"}
image(x,y,z2,col=color)
```

```{r ex2-2pltc, cache=TRUE, fig.cap="Interactive Plot", fig.pos="center"}
z2col  <- cut(z2, num_col)
plotid2 <- persp3d(x,y,z2, col=color[z2col])

rglwidget(elementId = "plot3drgl2") %>% playwidget(par3dinterpControl(fn, 0, 10, steps=25),
       step = 0.1, loop = TRUE, rate = 0.75) 
```

Lemma:


Lemma:


#### Ex 2.3

$H(x,y)=\frac{(x+1)(e^y-1)}{x+2e^y-1}$

```{r ex2-3, cache=TRUE}
# make vectorized function
H3<-function(x,y) mapply(function(x,y) ((x+1)*(exp(y)-1))/(x+2*exp(y)-1), x,y)

# create H-volume function
H3_vol<-mk_Hvol(fun=H3)

#check 2-increasing
#H3_vol(c(-1,1),c(1e-15,500))

x3<-seq(-1,1,length=50)
y3<-seq(0,15,length=50)
z3<-outer(x3,y3,FUN="H3")

#H3(x3[4],y3[2])==z3[4,2]
```

```{r ex2-3plta, cache=TRUE, fig.cap="Contour Plot", fig.pos="center"}
contour(x3,y3,z3)
```

```{r ex2-3pltb, cache=TRUE, fig.cap="Color Image", fig.pos="center"}
image(x3,y3,z3,col=color)
```

```{r ex2-3pltc, cache=TRUE, fig.cap="Interactive Plot", fig.pos="center"}
z3col  <- cut(z3, num_col)
plotid3<-persp3d(x3,y3,z3, col=color[z3col])

rglwidget(elementId = "plot3drgl3") %>% playwidget(par3dinterpControl(fn, 0, 10, steps=25),  step = 0.1, loop = TRUE, rate = 0.75)
```


Lemma:

### Copulas


### Exercises


### Sklar's Theorem

Definition:

#### Ex 2.4


Definition:

#### Ex 2.5

Theorem - Sklar's Theorem:

Lemma:

Lemma:

#### Ex 2.6

Definition:

#### Ex 2.7


### Copulas and Random Variables

### The Fréchet-Hoeffding Bounds for Joint Distribution Functions

### Survival Copulas

### Symmetry

### Order

### Random Variate Generation

### Multivariate Copulas


